Experience Report: Distributed Computing for the Common Man

### 2010-03-18 doc/icfp-2010/main.tex
*** empty log message ***
added more contents in introduction section; fixed some grammatical errors [KK]

This paper introduces a generic distributed computing library.
It was initially motivated by computing needs that exist in our own
research team. Our applications include large-scale deductive program
verification, which amounts to check the validity of a huge number of
logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, there is no
library that helps in exploiting such a computing infrastructure in
our favorite functional programming language. Hence we designed and
implemented such a library\footnote{This library is implemented in
  \Ocaml, but the implementation should be straightforward in any
  functional programming language.}, which is the subject of this paper.

what it is and what it is not
\begin{itemize}
\item it is not a library which does parallelization
\item it is a library for reliable distributed execution of
  parallelizable computation 
\end{itemize}

inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
programming, ironically) ;  however, there are differences:
\begin{itemize}
\item polymorphic, higher-order API
\item more generic (cores/network)
\item does not focus on association lists
\item we don't have data locality (future work)
\end{itemize}

what is the target audience/applications: 
for instance, use if automatic provers on thousands of verification
conditions, on computing infrastructures which can be
\begin{itemize}
\item a single machine with multiple cores, possibly remote
\item several machines, small or large in number, over a network
\end{itemize}

master and worker terminology introduced here

a few technical details about marshaling should be explained

>>>>

This paper introduces a generic distributed computing library. This
work was initially motivated by computing needs that exist in our own
research team. Our applications include large-scale deductive program
verification, which amounts to checking the validity of a large number
of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language. Hence we designed and implemented such a
library\footnote{This library is implemented in \Ocaml, but the
  implementation should be straightforward in any functional
  programming language.}, which is the subject of this paper.

The distributing computing library presented in this paper, is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows
execution of large-scale parallelizable computations, very relevant to
our application needs, (and presumably relevant to a variety of
real-world applications) over multiple cores in the same machine or
over a network of machines. It relieves the user from implementing
details of fault-tolerance, task distribution and parallelization. The
most striking feature of the library is its generality. The user needs
to change no more than a couple of lines in his code when he moves his
large-scale computation from a multi-core machine to a network of
machines. Further, the library provides exciting flexibilities as it
supports polymorphic functions, and rich features such as
fault-tolerance that become a cause for concern in any distributed
programming environment. It is worth noting that the library is not
targeted at applications running on server farms, crunching enormous
amounts of data. Neither it is limited to research labs possessing a
few number of powerful computing infrastructure. It serves a variety
of users and a wide spectrum of needs, from desktop PCs to network of
machines, and hence we find title \emph{Distributed Computing for the
  Common Man}, apt and appropriate.

% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}

% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}

% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}

\emph{Distributed Computing:} 
A typical distributed computing library, in its minimal implementation
should provide the following:
\begin{itemize}
\item A notion of a \of{task} which denotes the computation to be
  performed in a distributed manner
\item A set of machines or programs called \of{workers} that perform
  the tasks, producing results
\item A machine or a program called a \of{master} that is incharge
  of distributing the tasks among the workers and managing results
  produced by the workers
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance and efficient storage and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient serialization (or
marshalling) of data that can be reliably reconstructed when
transmitted over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment more often run on
different hardware architectures and make use of varying software
platforms. For example, a computer running on a hardware architecture
should be able to reliably reconstruct the data irrespective of
endianness. We carefully exploit the marshalling provided by \Ocaml\,
in particular the restrictions involved in marshalling functions, in
order to achieve maximally reliable transmission of messages over a
network.

The application domain of such a distributed computing library is
manyfold.  The most striking application, as also relevant to our
research endeavours, is in validating thousands of verification
conditions using automated theorem provers. Having noted this, it
should be apparent that the library finds use in efficient execution
of any large-scale expensive computation, that is parallelizable.

A couple of caveats are in order. (1) We do not focus on issues like
data-locality or distributed file system like Google's, rather we
assume a \unix\ Network File System (NFS). (2) Our focus is not
exactly on Google's Mapreduce paradigm~\cite{mapreduce} that works on
association lists, but we also provide an API similar to Google's that
works on association lists, just for satisfying popular interest.

Google's approach was also one of our inspirations to design and
implement a generic distributed computing library for functional
programming. Ironically, Google's approach itself was inspired by
functional programming primitives.

The paper is organized as follows. In Section~\ref{sec:API}, we
describe the main idea behind the generic API, and introduce
\of{master} and \of{worker} programs. In Section~\ref{sec:derived}, we
derive more interesting functions from the generic
API. Section~\ref{sec:implem} delves into the implementation details
of our library, while Section~\ref{sec:experiments} illustrates the
potential of the presented library through experimental evaluation. We
draw some conclusions in Section~\ref{sec:future} and outline our
future work.

% master and worker terminology introduced here
% a few technical details about marshaling should be explained
